---
title: "Adjusted competing risk curves simulation_study"
author: "Patrick van Hage"
output: 
  pdf_document:
    toc: TRUE
---

## Packages and supporting functions

### List of packages

Below is an overview of the packages used in this report, and a small description of their respective usage. 

```{r, message = F, warning = F}
library(AdjCuminc)        # Adjusted cumulative incidence curves        
library(openxlsx)         # Storing and loading output of simulation study
library(tidyverse)        # data wrangling and ggplot
library(ggpubr)           # Plot arrangements

# Adapted ggplot theme vankesteren/firatheme package 
theme_fira <- function(family = "sans") {
  ggplot2::`%+replace%`(
    ggplot2::theme_grey(base_size = 11.5, base_family = family),
    ggplot2::theme(
      # add padding to the plot
      plot.margin = grid::unit(rep(0.5, 4), "cm"),

      # remove the plot background and border
      plot.background = ggplot2::element_blank(),
      panel.background = ggplot2::element_blank(),
      panel.border = ggplot2::element_blank(),

      # make the legend and strip background transparent
      legend.background = ggplot2::element_rect(fill = "transparent",
                                                colour = NA),
      legend.key = ggplot2::element_rect(fill = "transparent",colour = NA),
      strip.background = ggplot2::element_rect(fill = "transparent",
                                               colour = NA),

      # add light, dotted major grid lines only
      panel.grid.major = ggplot2::element_line(linetype = "dotted",
                                               colour = "#454545",
                                               size = 0.3),
      panel.grid.minor = ggplot2::element_blank(),

      # remove the axis tick marks and hide axis lines
      axis.ticks = ggplot2::element_blank(),
      axis.line = ggplot2::element_line(color = "#454545", size = 0.3),

      # modify the bottom margins of the title and subtitle
      plot.title = ggplot2::element_text(size = 18, colour = "#454545",
                                         hjust = 0.5,
                                         margin = ggplot2::margin(b = 10)),
      plot.subtitle = ggplot2::element_text(size = 12, colour = "#454545",
                                            hjust = 0.5,
                                            margin = ggplot2::margin(b = 10)),

      # add padding to the caption
      plot.caption = ggplot2::element_text(size = 10, colour = "#454545",
                                           hjust = 1,
                                           margin = ggplot2::margin(t = 15)),

      # Adjust text size and axis title position
      axis.title = ggplot2::element_text(size = 13, colour = "#454545",
                                         hjust = 0.95),
      axis.text = ggplot2::element_text(size = 10, colour = "#212121"),
      legend.title = ggplot2::element_text(size = 12, colour = "#454545"),
      legend.text = ggplot2::element_text(size = 10, colour = "#454545"),
      strip.text = ggplot2::element_text(size = 10, colour = "#454545", 
                                         margin = ggplot2::margin(10, 10, 
                                                                  10, 10, 
                                                                  "pt")),
      axis.title.y = ggplot2::element_text(margin = unit(c(t = 0, 
                                                         r = 4,
                                                         b = 0, 
                                                         l = 0), "mm"),
                                           angle = 90)
    )
  )
}
```

## Visual evaluation of proposed adjustment methods

To visualize the effects of correction, a small simulation is performed. Confounding was introduced by correlating the set of discrete and continuous covariates $X_i$ with the allocated treatment group $Z_i$ ($Z_i = 0$ or $Z_i = 1$) via logistic regression. For reference, a "Target population" was added, where treatments were assigned randomly. 

### Validation of the true curve

Below we can demonstrate that the simulated data approximated the analytical true curve. 

```{r}
# Data simulation and biased sampling
times <- seq(0, 1.5, 0.05)
df <- simCSH(N = 10000, 
               k = 2, 
               theta_coef = c(-1, -0.5),
               beta_coef = list(c(1,  -1, 0.5), 
                                c(-1, 1, -0.5)),
               base_haz = c(2, -2))
df_true <- trueCSH(time = times, 
                     k = 2, 
                     theta_coef = c(-1, -0.5),
                     beta_coef = list(c(1,  -1, 0.5), 
                                      c(-1, 1, -0.5)),
                     base_haz = c(2, -2))

CumInc_df <- longCSH(survfit(Surv(obs_time, status) ~ treatment, data = df))
CumInc_true <- tidyr::gather(df_true, event, prob, 3:ncol(df_true))
colnames(CumInc_true)[2] <- "strata"
CumInc_true$strata <- factor(paste0("treatment=", CumInc_true$strata))

csh_CI <- rbind(CumInc_df, CumInc_true)
csh_CI$method <- c(rep("Reference population", nrow(CumInc_df)),
                   rep("True_curve", nrow(CumInc_true)))

# Produce plot
csh <- ggplot(csh_CI, aes(x = time, y = prob, col = method, lty = method)) + 
  facet_grid(strata ~ event) +
  geom_step(linewidth = 1) + 
  ylim(0, 1) + 
  labs(title = "Covariate adjustment on logistic treatment allocation", 
       col = "Method", 
       lty = "Method") 
csh
```

### Scenario 1. Adjustment of confounding by logistic treatment allocation 

In this scheme, confounding is introduced as a logistic association between the treatments and continuious and discrete covariates, under scenario 1. 

```{r}
# Data simulation and biased sampling
df <- simCSH(N = 10000, 
               k = 2, 
                    theta_coef = c(-1, -0.5),
                    beta_coef = list(c(1, -1, 0.5), 
                                     c(-1, 1, -0.5)),
               base_haz = c(0, 0))
df_conf1 <- confCSH(N = 4000, 
                    k = 2, 
                    theta_coef = c(-1, -0.5),
                    beta_coef = list(c(1, -1, 0.5), 
                                     c(-1, 1, -0.5)),
                    sample_coef = c(1, -1, 1),
                    control_ref = FALSE,
                    base_haz = c(0, 0))

# Construct weights
df_conf1$ipw<- adjIPW(treatment ~ x1 + x2, data = df_conf1)

# Perform adjustment and extract cumulative incidence curves
CumInc_df <- longCSH(survfit(Surv(obs_time, status) ~ treatment, data = df))
CumInc_Conf1 <- longCSH(survfit(Surv(obs_time, status) ~ treatment, data = df_conf1))
CumInc_IPW <- longCSH(survfit(Surv(obs_time, status) ~ treatment, 
                          data = df_conf1, weights = ipw))
CumInc_OR <- gather(adjOR(Hist(obs_time, status) ~ treatment + x1 + x2, 
              strata = "treatment", data = df_conf1), event, prob, 3:5)
CumInc_DR <- gather(adjDR(Hist(obs_time, status) ~ treatment + x1 + x2, 
              strata = "treatment", data = df_conf1), event, prob, 3:5)

csh_CI <- rbind(CumInc_df, CumInc_Conf1, CumInc_IPW, CumInc_OR, CumInc_DR)
csh_CI$method <- c(rep("Reference population", nrow(CumInc_df)),
                   rep("Confounded treatment allocation", nrow(CumInc_Conf1)),
                   rep("Inverse probability weights", nrow(CumInc_IPW)),
                   rep("Outcome regression modeling", nrow(CumInc_OR)),
                   rep("Doubly robust estimator", nrow(CumInc_DR)))
csh_CI$method <- factor(csh_CI$method, levels = c("Reference population",
                                                  "Confounded treatment allocation",
                                                  "Inverse probability weights",
                                                  "Outcome regression modeling",
                                                  "Doubly robust estimator"))

# Produce plot
csh1 <- ggplot(csh_CI, aes(x = time, y = prob, col = method, lty = method)) + 
  facet_grid(strata ~ event) +
  geom_step(linewidth = 1) + 
  ylim(0, 1) + 
  labs(title = "Covariate adjustment on logistic treatment allocation", 
       col = "Method", 
       lty = "Method") 
csh1
```

### Scenario 2. Adjustment of confounding by non-logistic treatment allocation 

In this scheme, confounding is introduced as a non-logistic association between the treatments and continuous and discrete covariates. Here, the composition of the treatment group is solely affected by the aforementioned association with the covariates. This can be achieved by setting `control_ref = TRUE`. The control group can be used as a reference group for the outcome regression modeling and doubly robust method. 

```{r}
# Data simulation and biased sampling
df <- simCSH(N = 10000, 
               k = 2, 
               theta_coef = c(-1, -0.5),
               beta_coef = list(c(1, -1, 0.5), 
                                c(-1, 1, -0.5)),
               base_haz = c(0, 0))
df_conf2 <- confCSH(N = 4000, 
                    k = 2, 
                    theta_coef = c(-1, -0.5),
                    beta_coef = list(c(1, -1, 0.5), 
                                     c(-1, 1, -0.5)),
                    sample_coef = c(1, -1, 1),
                    control_ref = TRUE,
                    base_haz = c(0, 0))

# Construct weights
df_conf2$ipw<- adjIPW(treatment ~ x1 + x2, data = df_conf2)

# Perform adjustment and extract cumulative incidence curves
CumInc_df <- longCSH(survfit(Surv(obs_time, status) ~ treatment, data = df))
CumInc_conf2 <- longCSH(survfit(Surv(obs_time, status) ~ treatment, data = df_conf2))
CumInc_IPW <- longCSH(survfit(Surv(obs_time, status) ~ treatment, 
                          data = df_conf2, weights = ipw))
CumInc_OR <- gather(adjOR(Hist(obs_time, status) ~ treatment + x1 + x2, 
              strata = "treatment", ref = "control", data = df_conf2), event, prob, 3:5)
CumInc_DR <- gather(adjDR(Hist(obs_time, status) ~ treatment + x1 + x2, 
              strata = "treatment", ref = "control", data = df_conf2), event, prob, 3:5)

csh_CI <- rbind(CumInc_df, CumInc_conf2, CumInc_IPW, CumInc_OR, CumInc_DR)
csh_CI$method <- c(rep("Reference population", nrow(CumInc_df)),
                   rep("Confounded treatment allocation", nrow(CumInc_conf2)),
                   rep("Inverse probability weights", nrow(CumInc_IPW)),
                   rep("Outcome regression modeling", nrow(CumInc_OR)),
                   rep("Doubly robust estimator", nrow(CumInc_DR)))
csh_CI$method <- factor(csh_CI$method, levels = c("Reference population",
                                                  "Confounded treatment allocation",
                                                  "Inverse probability weights",
                                                  "Outcome regression modeling",
                                                  "Doubly robust estimator"))

# Produce plot
csh2 <- ggplot(csh_CI, aes(x = time, y = prob, col = method, lty = method)) + 
  facet_grid(strata ~ event) +
  geom_step(lwd = 1) + 
  ylim(0, 1) + 
  labs(title = "Covariate adjustment on logistic treatment allocation", 
       col = "Method", 
       lty = "Method") 
csh2
```

### Scenario 3. Adjustment of confounding by logistic treatment allocation and non-linear outcome

In this scheme, confounding is introduced as a logistic association between the treatments and continuious and discrete covariates. However, the relation between the continuous covariate and outcome is incorporated as a piecewise linear effect of the continuous covariate on the baseline hazard, where if x1 > 1 returns the first element in `base_haz` for the baseline hazard, and the second element otherwise. Hence, if these two elements differ, a piecewise-linear effect is introduced on the baseline hazard. This effect may not accurately modeled by standard Cox proportional hazard regression methods. 

```{r}
# Data simulation and biased sampling
df <- simCSH(N = 10000, 
               k = 2,  
               theta_coef = c(-1, -0.5),
               beta_coef = list(c(1, -1, 0.5), 
                                c(-1, 1, -0.5)),
               base_haz = c(2, -2))
df_conf3 <- confCSH(N = 4000, 
                    k = 2, 
                    theta_coef = c(-1, -0.5),
                    beta_coef = list(c(1, -1, 0.5), 
                                     c(-1, 1, -0.5)),
                    sample_coef = c(1, -1, 1),
                    control_ref = FALSE,
                    base_haz = c(2, -2))

# Construct weights
df_conf3$ipw<- adjIPW(treatment ~ x1 + x2, data = df_conf3)

# Perform adjustment and extract cumulative incidence curves
CumInc_df <- longCSH(survfit(Surv(obs_time, status) ~ treatment, data = df))
CumInc_conf3 <- longCSH(survfit(Surv(obs_time, status) ~ treatment, data = df_conf3))
CumInc_IPW <- longCSH(survfit(Surv(obs_time, status) ~ treatment, 
                          data = df_conf3, weights = ipw))
CumInc_OR <- gather(adjOR(Hist(obs_time, status) ~ treatment + x1 + x2, 
              strata = "treatment", data = df_conf3), event, prob, 3:5)
CumInc_DR <- gather(adjDR(Hist(obs_time, status) ~ treatment + x1 + x2, 
              strata = "treatment", data = df_conf3), event, prob, 3:5)

csh_CI <- rbind(CumInc_df, CumInc_conf3, CumInc_IPW, CumInc_OR, CumInc_DR)
csh_CI$method <- c(rep("Reference population", nrow(CumInc_df)),
                   rep("Confounded treatment allocation", nrow(CumInc_conf3)),
                   rep("Inverse probability weights", nrow(CumInc_IPW)),
                   rep("Outcome regression modeling", nrow(CumInc_OR)),
                   rep("Doubly robust estimator", nrow(CumInc_DR)))
csh_CI$method <- factor(csh_CI$method, levels = c("Reference population",
                                                  "Confounded treatment allocation",
                                                  "Inverse probability weights",
                                                  "Outcome regression modeling",
                                                  "Doubly robust estimator"))

# Produce plot
csh3 <- ggplot(csh_CI, aes(x = time, y = prob, col = method, lty = method)) + 
  facet_grid(strata ~ event) +
  geom_step(lwd = 1) + 
  ylim(0, 1) + 
  labs(title = "Covariate adjustment on logistic treatment allocation with piecewise-linear covariate", 
       col = "Method", 
       lty = "Method") 
csh3
```

### Scenario 4. Adjustment of confounding by non-logistic treatment allocation and non-linear outcome

In this scheme, confounding is introduced as a non-logistic association between the treatments and continuous and discrete covariates. Here, the composition of the treatment group is solely affected by the aforementioned association with the covariates. In addition, the relation between the continuous covariate and outcome is incorporated as a piecewise linear effect of the continuous covariate on the baseline hazard, where if x1 > 1 returns the first element in `base_haz` for the baseline hazard, and the second element otherwise. Hence, if these two elements differ, a piecewise-linear effect is introduced on the baseline hazard. This effect may not accurately modeled by standard Cox proportional hazard regression methods. 

```{r}
# Data simulation and biased sampling
df <- simCSH(N = 10000, 
               k = 2,  
               theta_coef = c(-1, -0.5),
               beta_coef = list(c(1, -1, 0.5), 
                                c(-1, 1, -0.5)),
               base_haz = c(2, -2))
df_conf4 <- confCSH(N = 4000, 
                    k = 2, 
                    theta_coef = c(-1, -0.5),
                    beta_coef = list(c(1, -1, 0.5), 
                                     c(-1, 1, -0.5)),
                    sample_coef = c(1, -1, 1),
                    control_ref = TRUE,
                    base_haz = c(2, -2))

# Construct weights
df_conf4$ipw<- adjIPW(treatment ~ x1 + x2, data = df_conf4)

# Perform adjustment and extract cumulative incidence curves
CumInc_df <- longCSH(survfit(Surv(obs_time, status) ~ treatment, data = df))
CumInc_conf4 <- longCSH(survfit(Surv(obs_time, status) ~ treatment, data = df_conf4))
CumInc_IPW <- longCSH(survfit(Surv(obs_time, status) ~ treatment, 
                          data = df_conf4, weights = ipw))
CumInc_OR <- gather(adjOR(Hist(obs_time, status) ~ treatment + x1 + x2, 
              strata = "treatment", ref = "control", data = df_conf4), event, prob, 3:5)
CumInc_DR <- gather(adjDR(Hist(obs_time, status) ~ treatment + x1 + x2, 
              strata = "treatment", ref = "control", data = df_conf4), event, prob, 3:5)

csh_CI <- rbind(CumInc_df, CumInc_conf4, CumInc_IPW, CumInc_OR, CumInc_DR)
csh_CI$method <- c(rep("Reference population", nrow(CumInc_df)),
                   rep("Confounded treatment allocation", nrow(CumInc_conf4)),
                   rep("Inverse probability weights", nrow(CumInc_IPW)),
                   rep("Outcome regression modeling", nrow(CumInc_OR)),
                   rep("Doubly robust estimator", nrow(CumInc_DR)))
csh_CI$method <- factor(csh_CI$method, levels = c("Reference population",
                                                  "Confounded treatment allocation",
                                                  "Inverse probability weights",
                                                  "Outcome regression modeling",
                                                  "Doubly robust estimator"))

# Produce plot
csh4 <- ggplot(csh_CI, aes(x = time, y = prob, col = method, lty = method)) + 
  facet_grid(strata ~ event) +
  geom_step(lwd = 1) + 
  ylim(0, 1) + 
  labs(title = "Covariate adjustment on non-logistic treatment allocation with piecewise-linear covariate", 
       col = "Method", 
       lty = "Method") 
csh4
```

### Scenario 5. Violation of positivity assumption

In this scenario, the positivity assumption is violated by drastically increasing the odds of being sampled under Z = 1 for Stage 1 patients. As a result, the probability of being sampled under Z = 0 for stage 1 patients is close to 0, and the distribution of propensity scores has become assymmetric as a result. We can also observe that the propensity scores are heavily skewed for Z = 1. 

```{r}
# Data simulation and biased sampling
df <- simCSH(N = 10000, 
               k = 2, 
                    theta_coef = c(-1, -0.5),
                    beta_coef = list(c(1, -1, 0.5), 
                                     c(-1, 1, -0.5)),
               base_haz = c(0, 0))
df_conf5 <- confCSH(N = 4000, 
                    k = 2, 
                    theta_coef = c(-1, -0.5),
                    beta_coef = list(c(1, -1, 0.5), 
                                     c(-1, 1, -0.5)),
                    sample_coef = c(6, -1, 1),
                    control_ref = FALSE,
                    base_haz = c(0, 0))

# Construct weights
df_conf5$ipw<- adjIPW(treatment ~ x1 + x2, data = df_conf5)

# Perform adjustment and extract cumulative incidence curves
CumInc_df <- longCSH(survfit(Surv(obs_time, status) ~ treatment, data = df))
CumInc_conf5 <- longCSH(survfit(Surv(obs_time, status) ~ treatment, data = df_conf5))
CumInc_IPW <- longCSH(survfit(Surv(obs_time, status) ~ treatment, 
                          data = df_conf5, weights = ipw))
CumInc_OR <- gather(adjOR(Hist(obs_time, status) ~ treatment + x1 + x2, 
              strata = "treatment", data = df_conf5), event, prob, 3:5)
CumInc_DR <- gather(adjDR(Hist(obs_time, status) ~ treatment + x1 + x2, 
              strata = "treatment", data = df_conf5), event, prob, 3:5)

csh_CI <- rbind(CumInc_df, CumInc_conf5, CumInc_IPW, CumInc_OR, CumInc_DR)
csh_CI$method <- c(rep("Reference population", nrow(CumInc_df)),
                   rep("Confounded treatment allocation", nrow(CumInc_conf5)),
                   rep("Inverse probability weights", nrow(CumInc_IPW)),
                   rep("Outcome regression modeling", nrow(CumInc_OR)),
                   rep("Doubly robust estimator", nrow(CumInc_DR)))
csh_CI$method <- factor(csh_CI$method, levels = c("Reference population",
                                                  "Confounded treatment allocation",
                                                  "Inverse probability weights",
                                                  "Outcome regression modeling",
                                                  "Doubly robust estimator"))

# Produce plot
csh5 <- ggplot(csh_CI, aes(x = time, y = prob, col = method, lty = method)) + 
  facet_grid(strata ~ event) +
  geom_step(linewidth = 1) + 
  ylim(0, 1) + 
  labs(title = "Covariate adjustment on logistic treatment allocation", 
       col = "Method", 
       lty = "Method") 
csh5
```

Below, we can validate the effects on the distribution of propensity scores and covariates among the two treatment groups. 

```{r}
# Plot distribution of covariates
coef_plot <- ggplot(df_bias, aes(x = x2, 
                                 fill = treatment,
                                 col = treatment)) + 
  facet_grid(. ~ x1) +
  geom_histogram(aes(y = ifelse(after_stat(..count..) > 0,
                                after_stat(..count..), NA)),
                 data = ~ subset(., treatment == "control"), alpha = 0.5) +
  geom_histogram(aes(y = ifelse(after_stat(-..count..) < 0, 
                                after_stat(-..count..), NA)), 
                 data = ~ subset(., treatment == "treat"), alpha = 0.5) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
  annotate("segment", x=-Inf, xend=Inf, y=Inf, yend=Inf) +
  annotate("segment", x=Inf, xend=Inf, y=-Inf, yend=Inf) +
  theme_fira() +
  theme(legend.position = "top",
        axis.title.y = element_text(hjust = 0.5),
        axis.title.x = element_text(hjust = 0.5)) +
  labs(x = "Continuous covariate",
       y = "Frequency", 
       fill = "Exposure",
       col = "Exposure") +
  scale_y_continuous(limits = c(-200, 200), 
                     breaks = seq(-200, 200, 50), 
                     expand = c(0, 0)) +
  scale_x_continuous(limits = c(-4, 4), 
                     breaks = seq(-3, 3, 1), expand = c(0, 0)) 

# Plot distribution of propensity scores
score_plot <- ggplot(df_bias, aes(x = 1 / ipw, 
                                  fill = treatment,
                                  col = treatment)) + 
  geom_histogram(aes(y = ifelse(after_stat(..count..) > 0,
                                after_stat(..count..), NA)),
                 data = ~ subset(., treatment == "control"), alpha = 0.5) +
  geom_histogram(aes(y = ifelse(after_stat(-..count..) < 0, 
                                after_stat(-..count..), NA)), 
                 data = ~ subset(., treatment == "treat"), alpha = 0.5) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
  annotate("segment", x=-Inf, xend=Inf, y=Inf, yend=Inf) +
  annotate("segment", x=Inf, xend=Inf, y=-Inf, yend=Inf) +
  theme_fira() + 
  theme(legend.position = "top",
        axis.title.y = element_text(hjust = 0.5),
        axis.title.x = element_text(hjust = 0.5)) +
  labs(x = "Propensity score",
       y = "Frequency", 
       fill = "Exposure",
       col = "Exposure") +
  scale_y_continuous(limits = c(-1500, 1500), 
                     breaks = seq(-1500, 1500, 500)) 

# Arrange plots
plot_conf5a <- ggarrange(coef_plot, score_plot,
                         nrow = 1, ncol = 2, 
                         labels = c("A", "B"))
plot_conf5a
```

Save in high-quality format.

```{r}
ggsave("sim_output/sim_conf5a.tiff", 
       plot = plot_conf5a,
       units = "in", 
       width = 11.5, 
       height = 4, 
       dpi = 600, 
       compression = "lzw")
```

## Simulation study

### Scenario 1. Simulation of confounding by logistic treatment allocation 

A simulation study using $B = 1000$ iterations of data generation, adjustment, and subtraction from the true curve. In this scheme, confounding is introduced as a logistic association between the treatments and continuous and discrete covariates. Each plot shows the mean and interval between the 2.5th percentile and 97.5th percentile of the simulated differences with the true curve.  

```{r}
# start simulation
set.seed(1546732)

# Generate true curve for user-specified timeframe
times <- seq(0, 1.5, 0.05)
CumInc_true <- trueCSH(time = times, 
                    k = 2, 
                    theta_coef = c(-1, -0.5),
                    beta_coef = list(c(1, -1, 0.5), 
                                     c(-1, 1, -0.5)),
                    base_haz = c(0, 0))

B = 1000
n_methods = 4
deltas <- vector("list", B)
for(i in 1:B){
  
  # Simulate data
  df_conf1 <- confCSH(N = 4000, 
                      k = 2,
                      theta_coef = c(-1, -0.5),
                      beta_coef = list(c(1, -1, 0.5), 
                                       c(-1, 1, -0.5)),
                      sample_coef = c(1, -1, 1),
                      control_ref = FALSE,
                      base_haz = c(0, 0))
  sf_bias <- survfit(Surv(obs_time, status) ~ treatment, data = df_conf1)
  CumInc_bias <- summary(sf_bias, times = times, extend = T)$pstate
  
  # IPW correction
  df_conf1$ipw<- adjIPW(treatment ~ x1 + x2, data = df_conf1)
  sf_bias <- survfit(Surv(obs_time, status) ~ treatment, 
                     data = df_conf1, weights = ipw)
  CumInc_ipw <- summary(sf_bias, times = times, extend = T)$pstate
  CumInc_or <- adjOR(Hist(obs_time, status) ~ treatment + x1 + x2, 
                 strata = "treatment", data = df_conf1, times = times)[, 3:5]
  CumInc_dr <- adjDR(Hist(obs_time, status) ~ treatment + x1 + x2, 
                 strata = "treatment", data = df_conf1, times = times)[, 3:5]
  
  # Calculate difference w/ true curve
  tmp <- unname(cbind(rep(i, 2 * length(times) * n_methods),
                      rep(times, 
                          times = 2 * n_methods),
                      rep(c("control", "treat"), 
                          each = length(times), 
                          times = n_methods),
                      rep(c("1. Bias", 
                            "2. IPW", 
                            "3. OR", 
                            "4. DR"), 
                          each = 2 * length(times)),
                      rbind(CumInc_bias - CumInc_true[, 3:5],
                            CumInc_ipw - CumInc_true[, 3:5],
                            CumInc_or - CumInc_true[, 3:5],
                            CumInc_dr - CumInc_true[, 3:5])))
  names(tmp) <- c("iteration", 
                  "time", 
                  "treatment", 
                  "method",
                  "event_free", 
                  "event_1", 
                  "event_2")
  deltas[[i]] <- tmp
  
  # Denote iteration
  print(paste("iteration", i, "of", B))
}
output <- do.call("rbind", deltas)
output <- gather(output, event, delta, 5:7)
write.xlsx(output, "sim_output/conf1_t00_t15.xlsx")
```

```{r}
output <- read.xlsx("sim_output/conf1_t00_t15.xlsx")

# Summarize output simulation study
output_se <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) mean(x))
output_se$ymin <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) quantile(x, 0.025))$delta
output_se$ymax <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) quantile(x, 0.975))$delta

# Visualize results
sim_csh1 <- ggplot(output_se, 
                   aes(x = time, y = delta, col = treatment)) + 
  geom_line(lwd = 1) + 
  facet_grid(event ~ method) +
  geom_ribbon(data = output_se, 
              mapping = aes(ymin = ymin, ymax = ymax), 
              alpha = 0.2) +
  ylim(-0.3, 0.3) + 
  ggtitle("Covariate adjustment on logistic treatment allocation") 
sim_csh1
```

### Scenario 2. Simulation of confounding by non-logistic treatment allocation 

A small simulation study using $B = 100$ iterations of data generation, adjustment, and subtraction from the true curve. In this scheme, confounding is introduced as a non-logistic association between the treatments and continuous and discrete covariates. Here, the composition of the treatment group is solely affected by the aforementioned association with the covariates. The control group can be used as a reference group for the outcome regression modeling and doubly robust method. Each plot shows the mean and interval between the 2.5th percentile and 97.5th percentile of the simulated differences with the true curve. 

```{r}
# start simulation
set.seed(1546732)

# Generate true curve for user-specified timeframe
times <- seq(0, 1.5, 0.05)
CumInc_true <- trueCSH(time = times, 
                    k = 2, 
                    theta_coef = c(-1, -0.5),
                    beta_coef = list(c(1, -1, 0.5), 
                                     c(-1, 1, -0.5)), 
                    base_haz = c(0, 0))

B = 1000
n_methods = 4
deltas <- vector("list", B)
for(i in 1:B){
  
  # Simulate data
  df_conf2 <- confCSH(N = 4000, 
                      k = 2, 
                      theta_coef = c(-1, -0.5),
                      beta_coef = list(c(1, -1, 0.5), 
                                       c(-1, 1, -0.5)),
                      sample_coef = c(1, -1, 1),
                      control_ref = TRUE,
                      base_haz = c(0, 0))
  sf_bias <- survfit(Surv(obs_time, status) ~ treatment, data = df_conf2)
  CumInc_bias <- summary(sf_bias, times = times, extend = T)$pstate
  
  # Adjustment methods
  df_conf2$ipw<- adjIPW(treatment ~ x1 + x2, data = df_conf2)
  sf_bias <- survfit(Surv(obs_time, status) ~ treatment, 
                     data = df_conf2, weights = ipw)
  CumInc_ipw <- summary(sf_bias, times = times, extend = T)$pstate
  CumInc_or <- adjOR(Hist(obs_time, status) ~ treatment + x1 + x2, 
                strata = "treatment", ref = "control",
                data = df_conf2, times = times)[, 3:5]
  CumInc_dr <- adjDR(Hist(obs_time, status) ~ treatment + x1 + x2, 
                strata = "treatment", ref = "control", 
                data = df_conf2, times = times)[, 3:5]
  
  # Calculate difference w/ true curve
  tmp <- unname(cbind(rep(i, 2 * length(times) * n_methods),
                      rep(times, 
                          times = 2 * n_methods),
                      rep(c("control", "treat"), 
                          each = length(times), 
                          times = n_methods),
                      rep(c("1. Bias", 
                            "2. IPW", 
                            "3. OR", 
                            "4. DR"), 
                          each = 2 * length(times)),
                      rbind(CumInc_bias - CumInc_true[, 3:5],
                            CumInc_ipw - CumInc_true[, 3:5],
                            CumInc_or - CumInc_true[, 3:5],
                            CumInc_dr - CumInc_true[, 3:5])))
  names(tmp) <- c("iteration", 
                  "time", 
                  "treatment", 
                  "method",
                  "event_free", 
                  "event_1", 
                  "event_2")
  deltas[[i]] <- tmp
    
  # Denote iteration
  print(paste("iteration", i, "of", B))
}
output <- do.call("rbind", deltas)
output <- gather(output, event, delta, 5:7)
write.xlsx(output, "sim_output/conf2_t00_t15.xlsx")
```

```{r}
output <- read.xlsx("sim_output/conf2_t00_t15.xlsx")

# Summarize output simulation study
output_se <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) mean(x))
output_se$ymin <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) quantile(x, 0.025))$delta
output_se$ymax <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) quantile(x, 0.975))$delta

# Visualize results
sim_csh2 <- ggplot(output_se, 
                   aes(x = time, y = delta, col = treatment)) + 
  geom_line(lwd = 1) + 
  facet_grid(event ~ method) +
  geom_ribbon(data = output_se, 
              mapping = aes(ymin = ymin, ymax = ymax), 
              alpha = 0.2) +
  ylim(-0.3, 0.3) + 
  ggtitle("Covariate adjustment on non-logistic treatment allocation") 
sim_csh2
```

### Scenario 3. Simulation of confounding by logistic treatment allocation on non-linear outcome

A small simulation study using $B = 1000$ iterations of data generation, adjustment, and subtraction from the true curve. In this scheme, confounding is introduced as a logistic association between the treatments and continuious and discrete covariates. However, the relation between the continuous covariate and outcome is incorporated as a piecewise linear effect, which may not accurately modeled via Cox proportional hazard regression methods. Note that to specifically illustrate the effect of the non-linear relationship of the continuous covariate on the hazard, the discrete covariate is excluded from the simulation. Each plot shows the mean and interval between the 2.5th percentile and 97.5th percentile of the simulated differences with the true curve. 

```{r}
# start simulation
set.seed(1546732)

# Generate true curve for user-specified timeframe
times <- seq(0, 10, 0.25)
CumInc_true <- trueCSH(time = times, 
                    k = 2,
                    theta_coef = c(-1, -0.5),
                    beta_coef = list(c(1, -1, 0.5), 
                                     c(-1, 1, -0.5)),
                    base_haz = c(2, -2))

B = 1000
n_methods = 4
deltas <- vector("list", B)
for(i in 1:B){
  
  # Simulate data
  df_conf3 <- confCSH(N = 4000, 
                      k = 2, 
                      theta_coef = c(-1, -0.5),
                      beta_coef = list(c(1, -1, 0.5), 
                                       c(-1, 1, -0.5)),
                      sample_coef = c(1, -1, 1),
                      control_ref = FALSE,
                      base_haz = c(2, -2))
  sf_bias <- survfit(Surv(obs_time, status) ~ treatment, data = df_conf3)
  CumInc_bias <- summary(sf_bias, times = times, extend = T)$pstate
  
  # Adjustment methods
  df_conf3$ipw<- adjIPW(treatment ~ x1 + x2, data = df_conf3)
  sf_bias <- survfit(Surv(obs_time, status) ~ treatment, 
                     data = df_conf3, weights = ipw)
  CumInc_ipw <- summary(sf_bias, times = times, extend = T)$pstate
  CumInc_or <- adjOR(Hist(obs_time, status) ~ treatment + x1 + x2, 
                 strata = "treatment", data = df_conf3, times = times)[, 3:5]
  CumInc_dr <- adjDR(Hist(obs_time, status) ~ treatment + x1 + x2, 
                 strata = "treatment", data = df_conf3, times = times)[, 3:5]
  
  # Calculate difference w/ true curve
  tmp <- unname(cbind(rep(i, 2 * length(times) * n_methods),
                      rep(times, 
                          times = 2 * n_methods),
                      rep(c("control", "treat"), 
                          each = length(times), 
                          times = n_methods),
                      rep(c("1. Bias", 
                            "2. IPW", 
                            "3. OR", 
                            "4. DR"), 
                          each = 2 * length(times)),
                      rbind(CumInc_bias - CumInc_true[, 3:5],
                            CumInc_ipw - CumInc_true[, 3:5],
                            CumInc_or - CumInc_true[, 3:5],
                            CumInc_dr - CumInc_true[, 3:5])))
  names(tmp) <- c("iteration", 
                  "time", 
                  "treatment", 
                  "method",
                  "event_free", 
                  "event_1", 
                  "event_2")
  deltas[[i]] <- tmp
  
  # Denote iteration
  print(paste("iteration", i, "of", B))
}
output <- do.call("rbind", deltas)
output <- gather(output, event, delta, 5:7)
write.xlsx(output, "sim_output/conf3_t00_t10.xlsx")
```

```{r}
output <- read.xlsx("sim_output/conf3_t00_t10.xlsx")

# Summarize output simulation study
output_se <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) mean(x))
output_se$ymin <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) quantile(x, 0.025))$delta
output_se$ymax <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) quantile(x, 0.975))$delta

# Visualize results
sim_csh3 <- ggplot(output_se, 
                   aes(x = time, y = delta, col = treatment)) + 
  geom_line(lwd = 1) + 
  facet_grid(event ~ method) +
  geom_ribbon(data = output_se, 
              mapping = aes(ymin = ymin, ymax = ymax), 
              alpha = 0.2) +
  ylim(-0.3, 0.3) + 
  ggtitle("Covariate adjustment on logistic treatment allocation 
       with non-linear outcome effect of continuous covariate") 
sim_csh3
```

### Scenario 4. Simulation of confounding by non-logistic treatment allocation on non-linear outcome

A small simulation study using $B = 1000$ iterations of data generation, adjustment, and subtraction from the true curve. In this scheme, In this scheme, confounding is introduced as a non-logistic association between the treatments and continuous and discrete covariates. Here, the composition of the treatment group is solely affected by the aforementioned association with the covariates. In addition, the relation between the continuous covariate and outcome is incorporated as a piecewise linear effect, which may not accurately modeled via Cox proportional hazard regression methods. Note that to specifically illustrate the effect of the non-linear relationship of the continuous covariate on the hazard, the discrete covariate is excluded from the simulation. Each plot shows the mean and interval between the 2.5th percentile and 97.5th percentile of the simulated differences with the true curve. 

```{r}
# start simulation
set.seed(1546732)

# Generate true curve for user-specified timeframe
times <- seq(0, 10, 0.25)
CumInc_true <- trueCSH(time = times, 
                    k = 2,
                    theta_coef = c(-1, -0.5),
                    beta_coef = list(c(1, -1, 0.5), 
                                     c(-1, 1, -0.5)),
                    base_haz = c(2, -2))

B = 1000
n_methods = 4
deltas <- vector("list", B)
for(i in 1:B){
  
  # Simulate data
  df_conf4 <- confCSH(N = 4000, 
                      k = 2, 
                      theta_coef = c(-1, -0.5),
                      beta_coef = list(c(1, -1, 0.5), 
                                       c(-1, 1, -0.5)),
                      sample_coef = c(1, -1, 1),
                      control_ref = TRUE,
                      base_haz = c(2, -2))
  sf_bias <- survfit(Surv(obs_time, status) ~ treatment, data = df_conf4)
  CumInc_bias <- summary(sf_bias, times = times, extend = T)$pstate
  
  # Adjustment methods
  df_conf4$ipw<- adjIPW(treatment ~ x1 + x2, data = df_conf4)
  sf_bias <- survfit(Surv(obs_time, status) ~ treatment, 
                     data = df_conf4, weights = ipw)
  CumInc_ipw <- summary(sf_bias, times = times, extend = T)$pstate
  CumInc_or <- adjOR(Hist(obs_time, status) ~ treatment + x1 + x2, 
                strata = "treatment", ref = "control",
                data = df_conf4, times = times)[, 3:5]
  CumInc_dr <- adjDR(Hist(obs_time, status) ~ treatment + x1 + x2, 
                strata = "treatment", ref = "control", 
                data = df_conf4, times = times)[, 3:5]
  
  # Calculate difference w/ true curve
  tmp <- unname(cbind(rep(i, 2 * length(times) * n_methods),
                      rep(times, 
                          times = 2 * n_methods),
                      rep(c("control", "treat"), 
                          each = length(times), 
                          times = n_methods),
                      rep(c("1. Bias", 
                            "2. IPW", 
                            "3. OR", 
                            "4. DR"), 
                          each = 2 * length(times)),
                      rbind(CumInc_bias - CumInc_true[, 3:5],
                            CumInc_ipw - CumInc_true[, 3:5],
                            CumInc_or - CumInc_true[, 3:5],
                            CumInc_dr - CumInc_true[, 3:5])))
  names(tmp) <- c("iteration", 
                  "time", 
                  "treatment", 
                  "method",
                  "event_free", 
                  "event_1", 
                  "event_2")
  deltas[[i]] <- tmp
  
  # Denote iteration
  print(paste("iteration", i, "of", B))
}
output <- do.call("rbind", deltas)
output <- gather(output, event, delta, 5:7)
write.xlsx(output, "sim_output/conf4_t00_t10.xlsx")
```

```{r}
output <- read.xlsx("sim_output/conf4_t00_t10.xlsx")

# Summarize output simulation study
output_se <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) mean(x))
output_se$ymin <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) quantile(x, 0.025))$delta
output_se$ymax <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) quantile(x, 0.975))$delta

# Visualize results
sim_csh4 <- ggplot(output_se, 
                   aes(x = time, y = delta, col = treatment)) + 
  geom_line(lwd = 1) + 
  facet_grid(event ~ method) +
  geom_ribbon(data = output_se, 
              mapping = aes(ymin = ymin, ymax = ymax), 
              alpha = 0.2) +
  ylim(-0.3, 0.3) + 
  ggtitle("Covariate adjustment on non-logistic treatment allocation 
       with non-linear outcome effect of continuous covariate") 
sim_csh4
```

### Scenario 5. Violation of the positivity assumption

A simulation study using $B = 1000$ iterations of data generation, adjustment, and subtraction from the true curve. In this scheme, confounding is introduced as a logistic association between the treatments and continuous and discrete covariates. In addition, the positivity assumption has become violated by drastically increasing the odds of being sampled under Z = 1 for Stage 1 patients.

```{r}
# start simulation
set.seed(1546732)

# Generate true curve for user-specified timeframe
times <- seq(0, 1.5, 0.05)
CumInc_true <- trueCSH(time = times, 
                    k = 2, 
                    theta_coef = c(-1, -0.5),
                    beta_coef = list(c(1, -1, 0.5), 
                                     c(-1, 1, -0.5)),
                    base_haz = c(0, 0))

B = 1000
n_methods = 4
deltas <- vector("list", B)
for(i in 1:B){
  
  # Simulate data
  df_conf5 <- confCSH(N = 4000, 
                      k = 2,
                      theta_coef = c(-1, -0.5),
                      beta_coef = list(c(1, -1, 0.5), 
                                       c(-1, 1, -0.5)),
                      sample_coef = c(6, -1, 1),
                      control_ref = FALSE,
                      base_haz = c(0, 0))
  sf_bias <- survfit(Surv(obs_time, status) ~ treatment, data = df_conf5)
  CumInc_bias <- summary(sf_bias, times = times, extend = T)$pstate
  
  # IPW correction
  df_conf5$ipw<- adjIPW(treatment ~ x1 + x2, data = df_conf5)
  sf_bias <- survfit(Surv(obs_time, status) ~ treatment, 
                     data = df_conf5, weights = ipw)
  CumInc_ipw <- summary(sf_bias, times = times, extend = T)$pstate
  CumInc_or <- adjOR(Hist(obs_time, status) ~ treatment + x1 + x2, 
                 strata = "treatment", data = df_conf5, times = times)[, 3:5]
  CumInc_dr <- adjDR(Hist(obs_time, status) ~ treatment + x1 + x2, 
                 strata = "treatment", data = df_conf5, times = times)[, 3:5]
  
  # Calculate difference w/ true curve
  tmp <- unname(cbind(rep(i, 2 * length(times) * n_methods),
                      rep(times, 
                          times = 2 * n_methods),
                      rep(c("control", "treat"), 
                          each = length(times), 
                          times = n_methods),
                      rep(c("1. Bias", 
                            "2. IPW", 
                            "3. OR", 
                            "4. DR"), 
                          each = 2 * length(times)),
                      rbind(CumInc_bias - CumInc_true[, 3:5],
                            CumInc_ipw - CumInc_true[, 3:5],
                            CumInc_or - CumInc_true[, 3:5],
                            CumInc_dr - CumInc_true[, 3:5])))
  names(tmp) <- c("iteration", 
                  "time", 
                  "treatment", 
                  "method",
                  "event_free", 
                  "event_1", 
                  "event_2")
  deltas[[i]] <- tmp
  
  # Denote iteration
  print(paste("iteration", i, "of", B))
}
output <- do.call("rbind", deltas)
output <- gather(output, event, delta, 5:7)
write.xlsx(output, "sim_output/conf5_t00_t15.xlsx")
```

```{r}
output <- read.xlsx("sim_output/conf5_t00_t15.xlsx")

# Summarize output simulation study
output_se <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) mean(x))
output_se$ymin <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) quantile(x, 0.025))$delta
output_se$ymax <- aggregate(delta ~ time + treatment + method + event, 
                       data = output, FUN = function (x) quantile(x, 0.975))$delta

# Visualize results
sim_csh5 <- ggplot(output_se, 
                   aes(x = time, y = delta, col = treatment)) + 
  geom_line(lwd = 1) + 
  facet_grid(event ~ method) +
  geom_ribbon(data = output_se, 
              mapping = aes(ymin = ymin, ymax = ymax), 
              alpha = 0.2) +
  ylim(-0.3, 0.3) + 
  ggtitle("Covariate adjustment with violated positivity assumption") 
sim_csh5
```

### Bias and RMSE table

A table is added, indicating the bias and RMSE at the given time point within the range of simulated event times for each method of confounding. 

```{r}
sim_output <- list(read.xlsx("sim_output/conf1_t00_t15.xlsx"),
                   read.xlsx("sim_output/conf2_t00_t15.xlsx"),
                   read.xlsx("sim_output/conf3_t00_t10.xlsx"),
                   read.xlsx("sim_output/conf4_t00_t10.xlsx"))
ref_times <- c(0.8, 0.8, 5, 5) 
for(i in 1:length(sim_output)){
  sim_output[[i]] <- sim_output[[i]][sim_output[[i]]$time == ref_times[i], ]
  sim_output[[i]]$scheme <- rep(paste("conf", i), nrow(sim_output[[i]]))
}
sim_output <- do.call("rbind", sim_output)

# Summarize output simulation study (Bias & RMSE)
sim_bias <- aggregate(delta ~ scheme + method + treatment + event, 
                      data = sim_output, 
                      FUN = mean)
sim_bias$RMSE <- aggregate(delta ~ scheme + method + treatment + event, 
                      data = sim_output, 
                      FUN = function(x) sqrt(mean(x^2)))$delta
sim_bias[, 1:4] <- lapply(sim_bias[, 1:4], as.factor)

# Convert to table format
sim_table <- sim_bias %>% 
  nest(delta, RMSE, .key = "value_col") %>%
  spread(key = event, value = value_col) %>%
  unnest(event_1, event_2, event_free, .sep = "_")
as.data.frame(sim_table)
```

```{r}
write.xlsx(sim_table, "sim_output/output_table.xlsx")
```

### Visualization simulation study

Combine the plots from each simulation study into one format.

```{r}
sim_output <- list(read.xlsx("sim_output/conf1_t00_t15.xlsx"),
                   read.xlsx("sim_output/conf2_t00_t15.xlsx"),
                   read.xlsx("sim_output/conf3_t00_t10.xlsx"),
                   read.xlsx("sim_output/conf4_t00_t10.xlsx"))
for(i in 1:length(sim_output)){
  sim_output[[i]]$scheme <- rep(paste("conf", i), nrow(sim_output[[i]]))
}
sim_output <- do.call("rbind", sim_output)

# Summarize output simulation study
sim_summary <- aggregate(delta ~ scheme + time + treatment + method + event, 
                       data = sim_output, 
                       FUN = function (x) mean(x))
sim_summary$miny <- aggregate(delta ~ scheme + time + treatment + method + event, 
                       data = sim_output, 
                       FUN = function (x) quantile(x, 0.025))$delta
sim_summary$maxy <- aggregate(delta ~ scheme + time + treatment + method + event, 
                       data = sim_output, 
                       FUN = function (x) quantile(x, 0.975))$delta

# Adjust labels
sim_summary[, c(1, 3:5)] <- lapply(sim_summary[, c(1, 3:5)], as.factor)
levels(sim_summary$scheme) <- c("No misspecification", 
                                "Misspecified treatment", 
                                "Misspecified outcome",
                                "Both misspecified")
levels(sim_summary$method) <- c("Crude estimator", 
                                "Inverse probability \n weighting", 
                                "Outcome regression", 
                                "Doubly robust")
levels(sim_summary$treatment) <- c("Control", 
                                "Treatment")
levels(sim_summary$event) <- c("Event 1", 
                               "Event 2", 
                               "Event free")
```

```{r}
plot_conf1 <- ggplot(sim_summary[sim_summary$scheme == "No misspecification", ],
                     aes(x = time, y = delta, colour = treatment)) +
  facet_grid(event ~ method) + 
  geom_hline(yintercept = 0, lwd = 1, col = "#FFCC66", lty = 2) +
  geom_ribbon(aes(ymin = miny, ymax = maxy, x = time), alpha = 0.3) + 
  geom_line(lwd = 1) +
  ylim(-0.2, 0.2) + 
  labs(title = "No misspecification",
       x = "Time",
       y = "Difference from true curve",
       colour = "Exposure") +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(0, 0.4, 0.8, 1.2)) +
  theme_fira() + 
  theme(legend.position = "bottom",
        legend.margin = margin(t = -10))
plot_conf1
```

```{r}
plot_conf2 <- ggplot(sim_summary[sim_summary$scheme == "Misspecified treatment", ],
                     aes(x = time, y = delta, colour = treatment)) +
  facet_grid(event ~ method) + 
  geom_hline(yintercept = 0, lwd = 1, col = "#FFCC66", lty = 2) +
  geom_ribbon(aes(ymin = miny, ymax = maxy, x = time), alpha = 0.3) + 
  geom_line(lwd = 1) +
  ylim(-0.2, 0.2) + 
  labs(title = "Misspecified treatment model",
       x = "Time",
       y = "Difference from true curve",
       colour = "Exposure") +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(0, 0.4, 0.8, 1.2)) +
  theme_fira() +
  theme(legend.position = "bottom",
        legend.margin = margin(t = -10))
plot_conf2
```

```{r}
plot_conf3 <- ggplot(sim_summary[sim_summary$scheme == "Misspecified outcome", ],
                     aes(x = time, y = delta, colour = treatment)) +
  facet_grid(event ~ method) + 
  geom_hline(yintercept = 0, lwd = 1, col = "#FFCC66", lty = 2) +
  geom_ribbon(aes(ymin = miny, ymax = maxy, x = time), alpha = 0.3) + 
  geom_line(lwd = 1) +
  ylim(-0.2, 0.2) + 
  labs(title = "Misspecified outcome model",
       x = "Time",
       y = "Difference from true curve",
       colour = "Exposure") +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
  scale_x_continuous(expand = c(0, 0),
                     breaks = c(0, 2.5, 5.0, 7.5)) +
  theme_fira() +
  theme(legend.position = "bottom",
        legend.margin = margin(t = -10))
plot_conf3
```

```{r}
plot_conf4 <- ggplot(sim_summary[sim_summary$scheme == "Both misspecified", ],
                     aes(x = time, y = delta, colour = treatment)) +
  facet_grid(event ~ method) + 
  geom_hline(yintercept = 0, lwd = 1, col = "#FFCC66", lty = 2) +
  geom_ribbon(aes(ymin = miny, ymax = maxy, x = time), alpha = 0.3) + 
  geom_line(lwd = 1) +
  ylim(-0.2, 0.2) + 
  labs(title = "Both models misspecified",
       x = "Time",
       y = "Difference from true curve",
       colour = "Exposure") +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
  scale_x_continuous(expand = c(0, 0),
                     breaks = c(0, 2.5, 5.0, 7.5)) +
  theme_fira() +
  theme(legend.position = "bottom",
        legend.margin = margin(t = -10))
plot_conf4
```

Arrange and save the final format.

```{r}
plot_confs <- ggarrange(plot_conf1, 
                        plot_conf2,
                        plot_conf3,
                        plot_conf4,
                        labels = c("A", "B", "C", "D"), 
                        ncol = 1, nrow = 4)
plot_confs
```

```{r}
ggsave("sim_output/sim_plot.tiff", 
       plot = plot_confs,
       units = "in", 
       width = 11.5, 
       height = 16, 
       dpi = 600, 
       compression = "lzw")
```

### Visualization simulation of violated positivity assumption

Visualize the results from the simulation study with violated positivity. 

```{r}
sim_output <- read.xlsx("sim_output/conf5_t00_t15.xlsx")

# Summarize output simulation study
sim_summary <- aggregate(delta ~ time + treatment + method + event, 
                       data = sim_output, 
                       FUN = function (x) mean(x))
sim_summary$miny <- aggregate(delta ~ time + treatment + method + event, 
                       data = sim_output, 
                       FUN = function (x) quantile(x, 0.025))$delta
sim_summary$maxy <- aggregate(delta ~ time + treatment + method + event, 
                       data = sim_output, 
                       FUN = function (x) quantile(x, 0.975))$delta

# Adjust labels
sim_summary[, 2:4] <- lapply(sim_summary[, 2:4], as.factor)
levels(sim_summary$method) <- c("Crude estimator", 
                                "Inverse probability \n weighting", 
                                "Outcome regression", 
                                "Doubly robust")
levels(sim_summary$treatment) <- c("Control", 
                                "Treatment")
levels(sim_summary$event) <- c("Event 1", 
                               "Event 2", 
                               "Event free")
```

```{r}
plot_conf5 <- ggplot(sim_summary,
                     aes(x = time, y = delta, colour = treatment)) +
  facet_grid(event ~ method) + 
  geom_hline(yintercept = 0, lwd = 1, col = "#FFCC66", lty = 2) +
  geom_ribbon(aes(ymin = miny, ymax = maxy, x = time), alpha = 0.3) + 
  geom_line(lwd = 1) +
  labs(title = "Violated positivity assumption",
       x = "Time",
       y = "Difference from true curve",
       colour = "Exposure") +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(0, 0.4, 0.8, 1.2)) +
  scale_y_continuous(limits = c(-0.4, 0.4),
                     breaks = seq(-0.4, 0.4, 0.2)) +
  theme_fira() + 
  theme(legend.position = "bottom",
        legend.margin = margin(t = -10))
plot_conf5
```

```{r}
plot_conf5b <- ggarrange(plot_conf5,
                         nrow = 1, ncol = 1, 
                         labels = "C")
ggsave("sim_output/sim_conf5b.tiff", 
       plot = plot_conf5b,
       units = "in", 
       width = 11.5, 
       height = 4, 
       dpi = 600, 
       compression = "lzw")
```

In addition, the bias and RMSE table. 

```{r}
sim_output <- read.xlsx("sim_output/conf5_t00_t15.xlsx")
sim_output <- sim_output[sim_output$time == 0.8, ]
sim_output$scheme <- rep("conf 5", nrow(sim_output))

# Summarize output simulation study (Bias & RMSE)
sim_bias <- aggregate(delta ~ method + treatment + event, 
                      data = sim_output, 
                      FUN = mean)
sim_bias$RMSE <- aggregate(delta ~ method + treatment + event, 
                      data = sim_output, 
                      FUN = function(x) sqrt(mean(x^2)))$delta
sim_bias[, 1:3] <- lapply(sim_bias[, 1:3], as.factor)

# Convert to table format
sim_table <- sim_bias %>% 
  nest(delta, RMSE, .key = "value_col") %>%
  spread(key = event, value = value_col) %>%
  unnest(event_1, event_2, event_free, .sep = "_")
as.data.frame(sim_table)
```

```{r}
write.xlsx(sim_table, "sim_output/output_table_conf5.xlsx")
```
